---
title: "Chapter 1. Short-term rentals in British Columbia: Market overview"
output: pdf_document
date: '2022-05-16'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(slider)
library(lubridate)
qs::qload("output/data/ch_1.qsm", nthreads = 8)
col_palette <-
  c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

### Executive summary

**In 2021, there was an average of `r active_all_avg_2021` short-term rental listings active each day in the province of British Columbia (a decrease of `r active_change_pct` from 2019). The province's STR market is dominated by commercial operators---the top 10% of hosts earned `r host_top_10_pct` of all revenue, while the top 1% of hosts---just `r host_top_1_n` operators---earned `r host_top_1_pct` of revenue. The Covid pandemic caused a massive decline in STR activity. In total, from March 2020 through December 2021, we estimate that only `r covid_res_pct` of the reservations expected under the previous growth trend occurred. But this decline has been concentrated in the province's large urban regions, where reservations and prices have both plummeted. In smaller communities, reservations are down from their pre-pandemic trend, but nightly prices have been substantially higher, which suggests a redistribution of travel destinations from distant cities to nearby resort towns.**

### Active listings and host revenue

In 2021, there was an average of `r active_all_avg_2021` short-term rental listings active each day in the province of British Columbia. (Active listings are those which are not just displayed on a STR platform, but are either reserved or available for reservations, and they are the most straightforward way to assess the size of an STR market.) Setting aside the `r active_non_housing_2021` listings which were for hotels, B&Bs, and other non-housing properties, this leaves `r active_avg_2021` listings operated out of houses in BC. These listings were operated by `r hosts_avg_2021` hosts, who collectively earned `r rev_total_2021` in revenue (an average of `r rev_avg_2021` per listing, and a median of `r rev_med_2021` per listing).

These figures are all dramatically lower than the same ones for 2019---before the onset of the Covid-19 pandemic. In 2019, there were `r active_avg_2019` daily active listings operated out of houses in BC ( `r active_change_pct` higher than 2021). These listings were operated by `r hosts_avg_2019` hosts (`r hosts_change_pct` higher), who collectively earned `r rev_total_2019` in revenue (`r rev_change_pct` higher), with an average of `r rev_avg_2019` (`r rev_avg_change_pct` higher) and a median of `r rev_med_2019` (`r rev_med_change_pct` higher) per listing. (In what follows, we present figures from 2021---the most recent complete year, but one which is likely to be a historic low point for the STR market due to the Covid-19 pandemic. Where relevant, we also present figures from 2019---the last full year prior to the pandemic, and hence a plausible baseline for the STR market once the pandemic has subsided.)

Figure 1 displays the number of active listings per day across the province and in each of the "central city", "large urban", "small urban", "tourist destination" and "non-urban" sub-categories. The figure demonstrates that, while all types of community saw STR activity plunge at the beginning of the pandemic, large cities and their outskirts have still not seen anything resembling a recovery, while other communities saw STR activity rebound over the course of 2021. These patterns will be explored in more detail below.

```{r fig_1, echo = FALSE}

active_daily_tiers |> 
  group_by(tier) |> 
  mutate(n = slider::slide_dbl(n, mean, .before = 13)) |> 
  ungroup() |> 
  filter(date >= "2017-07-01", !is.na(tier)) |> 
  ggplot() +
  geom_line(aes(date, n, colour = tier)) +
  facet_wrap(~tier, scales = "free_y") +
  scale_color_brewer(palette = "Accent", guide = "none") +
  scale_y_continuous(labels = scales::comma) +
  ggtitle("Figure 1. Active STR listings per day") +
  theme_minimal()

```

### Home sharers and commercial operators

An important distinction in STR markets is between "home sharing" and "commercial operations". Home sharing occurs when the principal resident of a housing unit rents out part or all of that housing unit on a temporary basis. A family with a spare bedroom that they rent on Airbnb as a private-room listing is an example of home sharing, as is a condo unit whose occupant travels frequently for business, and rents out the entire unit when she is out of town. A commercial operation, by contrast, is an STR listing which is *not* located in the host's principal residence, and is operated in a more or less full-time fashion. Unlike home sharing arrangements, commercial STRs remove housing units from the long-term market, since each housing unit which is being operated as a full-time STR could instead be housing a long-term resident.

One way to distinguish between commercial operators and home sharers is to look at what type of listings are operated on STR platforms. STR listings can be entire-home or private-room, and the former is much more likely to be a home sharing arrangement than the latter. (Listings can also be shared-room, but these are very rare.) The overwhelming majority of the STR listings active in 2021 were entire homes (`r eh_pct_2021`), and this proportion increased noticeably during the pandemic (from `r eh_pct_2019` in 2019), presumably in part because non-commercial STR operators have become less willing to share their homes with strangers.

Another way to distinguish between home sharers and commercial operators is to examine the distribution of revenue among hosts. If revenue is evenly distributed among a large number of hosts earning modest amounts of money, that suggests that home sharing is common. If, by contrast, revenue is concentrated among a small number of high earners, that suggests that most STR activity is conducted by commercial operators, regardless of how many hosts are active on the platform. Across British Columbia, host revenue in 2021 was in fact highly concentrated. The top 10% of hosts earned `r host_top_10_pct` of all revenue, while the top 1% of hosts---just `r host_top_1_n` operators---earned `r host_top_1_pct` of revenue. By contrast, the median host earned `r rev_med_2021` in 2021. This pattern suggests that BC has a large number of hosts who are are casual home sharers, but that the STR market is dominated by a much smaller number of commercial operators. Figure 2 shows the distribution of host revenue for the province and each of the sub-categories. Revenue is more highly concentrated in smaller resort destinations (implying a high dominance of commercial operations), while in non-urban, non-resort communities revenue is somewhat more equitably distributed (implying more presence of genuine home sharing).

```{r fig_2, echo = FALSE}

host_deciles |> 
  ggplot(aes(position, value, group = decile, fill = decile)) +
  geom_area(colour = "white", lwd = 1.2) +
  facet_wrap(~ tier, nrow = 2) +
  scale_y_continuous(name = "Host decile", label = scales::label_percent(1),
                     breaks = seq(0, 1, by = 0.1), limits = c(0, 1),
                     sec.axis = sec_axis(~., 
                                         name = "% of total revenue",
                                         labels = derive(), 
                                         breaks = derive())) +
  scale_fill_gradientn(colours = revenue_colour) +
  ggtitle("Figure 2. Host revenue distribution (2021)") +
  theme_void() +
  theme(legend.position = "none",
        # text = element_text(family = "Futura"),
        axis.text.y = element_text(hjust = 1),
        axis.title.y.left = element_text(
          angle = 90, margin = margin(0, 10, 0, 0)),
        axis.title.y.right = element_text(
          angle = 270, margin = margin(0, 0, 0, 10)),
        axis.title.x = element_blank(), 
        axis.text.x = element_blank(),
        plot.background = element_rect(fill = "white", colour = "transparent"))

```

Finally, one simple method of identifying commercial STR operators is to identify listings operated by hosts who have multiple listings in operation simultaneously. These "multilistings" by definition cannot be home sharing arrangements, since the host cannot have multiple principal residences. We consider entire-homes to be "multilistings" if they are operated by hosts who are simultaneously operating other entire-home listings. We define private-room multilistings as cases where a host has three or more private-room listings operating on the same day. Since nearly all entire-home listings have three or fewer bedrooms, there will be extremely few cases where a host operating three private-room STR listings in a dwelling unit has not converted the entire unit into a dedicated STR.

In 2021, `r ml_pct_2021` of active listings in British Columbia were multilistings, earning `r ml_rev_pct_2021` of total host revenue. Figure 3 shows the share of all active listings which are multilistings over time. Multilistings have been growing steadily since 2017, both in terms of listings and revenue percentage, and they are much more common in resort communities and large cities than in smaller, non-resort communities.

```{r fig_3, echo = FALSE}

ml_active |> 
  group_by(tier) |> 
  mutate(n = slider::slide_dbl(pct, mean, .before = 13)) |> 
  ungroup() |> 
  filter(date >= "2017-07-01", !is.na(tier)) |> 
  ggplot() +
  geom_line(aes(date, n, colour = tier)) +
  facet_wrap(~tier) +
  scale_color_brewer(palette = "Accent", guide = "none") +
  scale_y_continuous(labels = scales::percent) +
  ggtitle("Figure 3. Multilistings as a share of active daily listings") +
  theme_minimal()

```

By any of these metrics, BC's STR market is increasingly dominated by commercial operators at the expense of home sharers, who have been pushed to the margins of the market. Commercial operators are particularly common in resort communities and large cities, but there is no part of the province where they are not dominant.

### Growth trends: pre- and post-Covid

Until early 2020, the number of active STR listings in British Columbia was steadily increasing (with the exception of Vancouver, where the 2018 introduction of the City's STR regulations caused a large one-time drop in the number of active listings). Figure 4 shows the change in active listings relative to one year earlier, which is a convenient way to remove seasonal variation to identify underlying growth trends. The figure indicates that, from 2018 through the beginning of the pandemic, active listings were growing in all different community types. The pandemic halted BC's STR market growth, with active listings collapsing in March 2020. But, while listing growth again turned positive in smaller communities as of early 2021, active listings continued to shrink until the end of 2021 in BC's largest cities.

```{r fig_4, echo = FALSE}

daily_variation |> 
  group_by(tier) |> 
  mutate(n = slide_dbl(n, mean, .before = 20, .complete = TRUE)) |> 
  ungroup() |> 
  ggplot(aes(date, n, colour = tier)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +
  geom_line(lwd = 1, na.rm = TRUE) +
  scale_x_date(name = NULL) +
  scale_y_continuous(name = NULL, limits = c(-1, NA), 
                     labels = scales::percent) +
  scale_color_brewer(palette = "Accent", guide = "none") +
  facet_wrap(~tier) +
  theme_minimal() +
  theme(legend.position = "none", panel.grid.minor.x = element_blank())

```

Overall, the year-over-year change in average active listings from 2017 to 2018 was `r active_change_pct_2018`, and the year-over-year change from 2018 to 2019 was `r active_change_pct_2019`. Because of the Covid-19 pandemic, the year-over-year change in active daily listings was `r active_change_pct_2020` in 2020 and `r active_change_pct_2021` in 2021.

#### Trend analysis: reservations

The pandemic had a massive impact on STR listings and revenue throughout the province, and trend analysis allows for a more precise estimate of how this impact varied from location to location. Figure 5 provides a closer look at daily reservations since 2019, comparing the actual trajectory of reservations during the pandemic with what the trajectory of reservations would have been expected to be, based on previous growth in the overall provincial STR market and in each of the sub-markets, but in the absence of the pandemic. (To do this, we use seasonal decomposition to identify the regular seasonal fluctuations in STR activity and separate them from the underlying patterns of growth or decline.)

```{r fig_5, echo = FALSE, warning = FALSE, fig.width = 9}

reservations_and_prices |> 
  group_by(tier) |> 
  mutate(res = slide_dbl(res, mean, .before = 6, .complete = TRUE)) |> 
  ungroup() |> 
  mutate(res_trend = if_else(date <= "2020-03-14", NA_real_, res_trend)) |> 
  select(tier, date, res, res_trend) |> 
  pivot_longer(-c(tier, date)) |> 
  filter(!is.na(value)) |> 
  mutate(label = case_when(
    tier == "All" & date == "2019-07-05" & name == "res" ~ "Actual reservations", 
    tier == "All" & date == "2020-08-05" & name == "res_trend" ~ "Expected reservations",
    TRUE ~ NA_character_)) |> 
  ggplot() +
  geom_ribbon(aes(x = date, ymin = res, ymax = res_trend, group = 1),
              data = {reservations_and_prices |> 
                  group_by(tier) |> 
                  mutate(res = slider::slide_dbl(res, mean, .before = 6, .complete = TRUE)) |> 
                  ungroup() |> 
                  mutate(res_trend = if_else(date <= "2020-03-14", NA_real_, res_trend))
                }, fill = col_palette[3],
              alpha = 0.3) +
  geom_line(aes(date, value, color = name), lwd = 1) +
  geom_label(aes(date, value, label = label, color = name), size = 3) +
  scale_x_date(name = NULL, limits = as.Date(c("2018-01-01", NA))) +
  scale_y_continuous(name = NULL, limits = c(0, NA), 
                     label = scales::comma) +
  scale_color_manual(name = NULL, 
                     labels = c("Actual reservations", "Expected reservations"), 
                     values = col_palette[c(5, 1)]) +
  facet_wrap(~tier, scales = "free_y") +
  theme_minimal() +
  theme(legend.position = "none",
        panel.grid.minor.x = element_blank(),
        plot.background = element_rect(fill = "white", colour = "transparent"))

```

As Figure 5 suggests, the impact of the pandemic on STR reservations has been highly uneven between different types of communities in the province. In total, from March 2020 through December 2021, we estimate that there have been `r covid_res_dif` fewer STR nights reserved than would normally have been expected to occur. The `r covid_res_total` total nights reserved in this time period is only `r covid_res_pct` of the `r covid_res_trend` nights total that would represent the previous growth trend. However, an enormous amount of these "missing" reservations are concentrated in central cities. Here, actual reservations were only `r covid_cc_res_pct` of what the pre-pandemic trend would have predicted. By contrast, in resort communities, actual reservations were `r covid_res_res_pct` of the pre-pandemic trend---only two thirds the decline experienced in central cities.

#### Trend analysis: nightly prices

Unlike reservations, which fell consistently---if at slightly different rates---across the province during the pandemic, nightly prices have behaved very differently in different community types (Figure 6). In central cities and the rest of the large urban regions, prices fell precipitously in 2020. In combination with the steep drop in reservations during this same time period, this unambiguously describes a collapse in STR demand in the big cities, as long-distance travel which normally supplies the temporary accommodation demand dried up. Throughout 2021, prices generally remained below their pre-pandemic trend, albeit at somewhat higher levels than 2020.

In smaller communities and resort towns, however, the pattern in 2020 was nearly the opposite. Compared to the pre-pandemic trend, in many of these locations prices *increased* substantially. This occurred at the same time as actual reservations declined in these locations. This combination suggests a spike in demand for non-urban STRs in the early pandemic---consistent with the idea that leisure travelers cancelled their long-distance vacations and substituted local vacations instead---but also that the actual supply of STRs in these locations was constrained, resulting in higher prices rather than increased reservations. Prices have remained much higher than the pre-pandemic trend in these non-urban locations throughout 2021.

```{r fig_6, echo = FALSE, warning = FALSE, fig.width = 9}

reservations_and_prices |> 
  group_by(tier) |> 
  mutate(price = slide_dbl(price, mean, .before = 6, .complete = TRUE)) |> 
  ungroup() |> 
  select(tier, date, price, price_trend) |> 
  pivot_longer(-c(tier, date)) |> 
  filter(!is.na(value)) |> 
  mutate(label = case_when(
    tier == "All" & date == "2019-07-05" & name == "price" ~ "Actual price", 
    tier == "All" & date == "2020-08-05" & name == "price_trend" ~ "Expected price",
    TRUE ~ NA_character_)) |> 
  ggplot() +
  geom_ribbon(aes(x = date, ymin = price, ymax = price_trend, group = 1),
              data = {reservations_and_prices |> 
                  group_by(tier) |> 
                  mutate(price = slider::slide_dbl(price, mean, .before = 6, 
                                                   .complete = TRUE)) |> 
                  ungroup()}, fill = col_palette[3],
              alpha = 0.3) +
  geom_line(aes(date, value, color = name), lwd = 1) +
  geom_label(aes(date, value, label = label, color = name), size = 3) +
  scale_x_date(name = NULL, limits = as.Date(c("2018-01-01", NA))) +
  scale_y_continuous(name = NULL, limits = c(0, NA), 
                     label = scales::comma) +
  scale_color_manual(name = NULL, 
                     labels = c("Actual price", "Expected price"), 
                     values = col_palette[c(5, 1)]) +
  facet_wrap(~tier, scales = "free_y") +
  theme_minimal() +
  theme(legend.position = "none",
        panel.grid.minor.x = element_blank(),
        plot.background = element_rect(fill = "white", colour = "transparent"))

```

### Conclusions

British Columbia's STR market over the last several years can be characterized by the following stylized facts:

1.  Commercial operators dominate the STR market province-wide, and in each community type, although they are particularly dominant in large cities and resort towns. In general, the top 10% of STR hosts earn a majority of all the STR revenue in the province.
2.  During the pandemic there has been a strong divergence between big cities and smaller communities. The former saw their STR markets collapse, and they have only barely begun to recover. The latter saw declines in STR bookings, but increases in nightly prices.
3.  The patterns are consistent with a redistribution of travel destinations from distant cities to nearby resort towns.
