---
title: "Chapter 2. The impact of short-term rentals on housing availability and affordability in British Columbia"
output: pdf_document
date: '2022-05-25'
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)

library(tidyverse)
library(slider)
qs::qload("output/data/ch_2.qsm")
col_palette <-
  c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

```

## Executive summary

**At the end of 2021, short-term rentals were taking `r housing_loss_2021` housing units off of BC's long-term market. This was a `r housing_loss_decline_pct_2019_2021` decline compared to 2019 (`r housing_loss_2019`), which is a much smaller decrease than the overall STR market saw; province-wide, commercial operators have weathered the pandemic much better than home sharers. We develop a regression model to measure the impact of commercial STRs on average rents in BC, and find that an increase of one dedicated STR per 100 rental units predicts that average rents in the neighbourhood will be `r model_iv_coef_dollar` higher. From 2016-2021, we estimate that, if there had been no commercial STRs, BC renters would have saved `r overpaid_2016_2021` in rent. In fact, during the 2017-2019 pre-pandemic period, the proliferation of dedicated STRs across BC accounts for fully `r str_incr_2017_2019` of the increases in rents which BC households were subjected to. The average BC neighbourhood in a medium or large city saw a `r rent_month_2017_2019` increase in monthly rent each of these years, of which `r str_incr_month_2017_2019` can be attributed to the growth of dedicated STRs. In 2020, when the number of commercial STRs plummeted because of the pandemic, our estimate is that 2020 rents increased `r str_incr_2020` less than they otherwise would have, because of the decrease in commercial STRs that year. If the province's STR market returns to its pre-pandemic trajectory by the end of 2023, the loss of housing to STRs will be expected to drive up average monthly rents in medium and large cities in the province by `r rent_inc_monthly_2021_2023` above their levels at the end of 2021.**

## Introduction

BC's housing market has been under considerable stress in the past years, with housing prices and rents rising, and rental vacancy rates falling. These are symptoms of a market where the supply of housing is insufficient to meet demand. One possible explanation for both the insufficient supply and elevated demand for housing in BC is the growth in short-term rentals. Tourists are now able to compete with residents for housing---adding demand to the local housing market---while landlords are now able to shift their properties out of the conventional housing market to become dedicated STRs---reducing the supply of conventional housing. Research has found that renting a housing unit on the STR market frequently offers landlords greater potential revenue than conventional leases (Wachsmuth & Weisler 2018), especially in transit-accessible neighborhoods (Deboosere et al. 2019). Multiple studies have also found that Airbnb and other STR platforms increase housing costs (Barron, Kung, & Proserpio 2017; Horn & Merante 2017; Garcia-Lopez et al. 2019).

This chapter combines our unique dataset of STR activity in BC with housing data from the 2016 Canadian Census and the Canadian Mortgage and Housing Corporation's Rental Market Survey to determine the impact of commercially-operated short-term rentals on housing availability, measured through the number of housing units removed form the long-term market by STRs, and housing affordability, measured through average market rents.

## STR-induced housing loss

One of the major considerations when gauging the impacts of short-term rentals (STRs) on a city is the extent to which STRs are removing long-term housing from the market. This process can occur either directly, where tenants of a unit are evicted or not replaced at the end of a lease and the unit is converted to a STR, or indirectly by absorbing new construction or investment properties which otherwise would have gone onto the long-term market. To obtain the exact number of units that have been occupied as STRs, landlords or units would need to be individually surveyed, which is infeasible because STR hosts are mostly anonymous on major STR platforms such as Airbnb and Vrbo. Instead, we use the daily activity of listings, alongside structural characteristics such as listing type and location, to estimate which listings are operating as dedicated STRs and are therefore not available as conventional long-term housing.

*Frequently Rented Entire-Home (FREH) listings:* The number of frequently-rented units is one way to estimate STR-induced housing loss. If a STR is available for reservations the majority of the year and receives many bookings, it is reasonable to assume that it is not serving as an individual's principal residence at the same time. Along these lines, we define frequently rented entire-home (FREH) listings as entire-home listings which were available on Airbnb or Vrbo the majority of the year (at least 183 nights) and were booked a minimum of 90 nights. We then apply a statistical model (described in the appendix) to the FREH data in order to generate an estimate of FREH activity based on three months of listing activity. This allows us to detect listings which are operating in a full-time manner but have not yet been listed for an entire year, and allows us to account for relatively short-term changes in market conditions.

*Ghost hostels:* In addition to FREH listings, it is possible that entire housing units have been subdivided into multiple private-room listings, each of which appearing to be a spare bedroom or the like, while actually collectively representing an apartment removed from the long-term housing market. We call these clusters of private- room listings "ghost hostels", building on the advocacy group Fairbnb.ca's term "ghost hotels" --- multiple FREH listings located in a single building, collectively serving as de facto hotels instead of long-term housing (Wieditz 2017). We detect ghost hostels by finding clusters of three or more private-room listings operated by a single host, whose reported locations are close enough to each other that they are likely to have originated in the same actual housing unit. (Airbnb and Vrbo obfuscate listing locations by shifting them randomly up to 200 m.)

At the end of 2021, there were `r freh_2021` FREH listings in British Columbia, and `r gh_units_2021` more housing units which were operating as ghost hostels. (We refer to these listings collectively as "dedicated STRs".) In total, therefore, short-term rentals were taking `r housing_loss_2021` housing units off of BC's long-term market at the end of the year (Figure 7). In the absence of commercial STRs, in other words, there would have been `r housing_loss_2021` more homes available for BC residents to live in. Notably, while the number of active daily listings was `r active_decline_pct_2019_2021` lower in 2021 than in 2019, the number of housing units which STRs took off of BC's housing market decreased at only about half the rate. There were `r housing_loss_decline_pct_2019_2021` fewer dedicated STRs in 2021 (`r housing_loss_2021`) than in 2019 (`r housing_loss_2019`); province-wide, commercial operators have weathered the pandemic much better than home sharers.

```{r fig_7, fig.width = 9}

housing_loss |> 
  ggplot(aes(date, `Housing units`, fill = `Listing type`)) +
  geom_col(lwd = 0) +
  scale_fill_manual(values = col_palette[c(2, 4)]) +
  scale_x_date(name = NULL, limits = c(as.Date("2017-07-01"), NA)) +
  scale_y_continuous(name = NULL, label = scales::comma) +
  facet_wrap(~tier, scales = "free_y") +
  theme_minimal() +
  theme(legend.position = "bottom", 
        panel.grid.minor.x = element_blank())

```

As with other patterns related to STR activity during the pandemic, the large cities saw much more severe drops in dedicated STRs than smaller communities and resort towns did. The latter had nearly recovered to their pre-pandemic levels of dedicated STRs by the beginning of 2022, while the former were only beginning to emerge from the pandemic's negative demand shock. These facts further confirm the scenario identified above: that long-distance leisure trips were diverted to local destinations during the pandemic.

### Trend analysis: housing loss

Using trend analysis, it is possible to estimate how much STR-induced housing loss would have been expected to have occurred in the absence of the pandemic---this counterfactual scenario establishes an important baseline for a post-pandemic housing landscape, in which short-term rentals return to their previous growth trend. Figure 8 displays the shortfall between actual STR-induced housing loss and the housing loss which would have been expected in the absence of the pandemic. In the central cities and other areas of the largest urban areas of the province, there are several thousand fewer dedicated STRs than what pre-pandemic growth trends would have predicted. In particular, at the end of 2021, there were only `r housing_loss_cc_end_2021` housing units operating as dedicated STRs in Vancouver, Victoria, and Abbotsford, while the previous trend would have predicted `r housing_loss_cc_trend_end_2021`---`r housing_loss_cc_dif_pct_end_2021` more. By contrast, resort towns and non-urban areas had very little short fall between expected and actual dedicated STRs by the end of 2021. These areas have seen relatively robust and growing tourist accommodation demand from late 2020 onward.

```{r fig_8, fig.width = 9}

housing_loss_daily_model |> 
  group_by(tier) |>
  mutate(units_trend = slide_dbl(units_trend, mean, .before = 6, .complete = TRUE)) |>
  ungroup() |>
  select(tier, date, units, units_trend) |> 
  pivot_longer(-c(tier, date)) |> 
  filter(!is.na(value)) |> 
  mutate(label = case_when(
    tier == "All" & date == "2019-05-05" & name == "units" ~ "Actual housing loss", 
    tier == "All" & date == "2020-03-07" & name == "units_trend" ~ "Expected housing loss",
    TRUE ~ NA_character_)) |> 
  ggplot() +
  geom_ribbon(aes(x = date, ymin = units, ymax = units_trend, group = 1),
              data = {
                housing_loss_daily_model |> 
                  group_by(tier) |>
                  mutate(units_trend = slider::slide_dbl(units_trend, mean, .before = 6, .complete = TRUE)) |>
                  ungroup()}, fill = col_palette[3], alpha = 0.3) +
  geom_line(aes(date, value, color = name), lwd = 1) +
  geom_label(aes(date, value, label = label, color = name), size = 3) +
  scale_x_date(name = NULL, limits = as.Date(c("2018-01-01", NA))) +
  scale_y_continuous(name = NULL, limits = c(0, NA), 
                     label = scales::comma) +
  scale_color_manual(name = NULL, 
                     labels = c("Actual STR housing loss", 
                                "Expected STR housing loss"), 
                     values = col_palette[c(5, 1)]) +
  facet_wrap(~tier, scales = "free_y") +
  theme_minimal() +
  theme(legend.position = "none",
        panel.grid.minor.x = element_blank(),
        plot.background = element_rect(fill = "white", colour = "transparent"))

```

## The impact of dedicated STRs on residential rents in BC

While no empirical research exists in a Canadian context to evaluate the impact of STRs on housing prices or rents, Barron et al. (2020) answered these questions through an examination of every US Airbnb listing between 2012 and 2016. This study found that a 1% growth in Airbnb listings in a location predicts a 0.018% increase in monthly rents and a 0.026% increase in house prices. While these numbers might appear small, they occurred in the context of STR growth rates which were quite high; the authors find that the growth of Airbnb was responsible for *one fifth of all rent growth and one seventh of all housing-price growth* in the United States during the study period. This finding, while not directly applicable to British Columbia, establishes an *a priori* rationale for testing the hypothesis that the growth of STRs in BC has had a similar impact on housing affordability.

### A regression model of dedicated STRs and residential rent

Rents have been increasing steadily across all community types in British Columbia since 2016, when our data series begins. Figure 9 shows the change in average monthly rents in each neighbourhood in the province from 2016 to 2021. Across 2016-2021, average monthly rents were highest in the central cities (`r rent_tier[["CC"]]`) and surrounding areas (`r rent_tier[["CMA"]]`) of the largest urban regions, followed by resort towns and non-urban areas (`r rent_tier[["RES/NU"]]`), and then mid-sized cities (`r rent_tier[["CA"]]`). (The Canada Mortgage and Housing Corporation only administers the Rental Market Survey in mid-sized and large urban regions---"census agglomerations" and "census metropolitan areas" respectively---so we lack data for many smaller communities which are nevertheless important tourist destinations. Most notably, Whistler is excluded from this analysis, as are small communities such as Tofino and Ucluelet. As a result, in this section we merge the "resort town" and "non-urban" communities into a single category.)

```{r fig_9, fig.width = 9}

cmhc$rent |> 
  filter(!is.na(tier)) |> 
  mutate(tier = "All") |> 
  bind_rows(cmhc$rent) |> 
  select(tier, year, neighbourhood, total) |> 
  filter(!is.na(tier)) |> 
  mutate(tier = if_else(tier %in% c("NU", "RES"), "RES/NU", tier)) |> 
  ggplot(aes(year, total, colour = tier)) +
  geom_line(aes(group = neighbourhood), alpha = 0.25) +
  # geom_jitter(width = 0.15, height = 0, alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE) +
  scale_x_continuous(name = NULL) +
  scale_y_continuous(name = NULL, labels = scales::dollar) +
  scale_color_brewer(palette = "Accent", guide = "none") +
  facet_wrap(~tier) +
  ggtitle("Average rents for CMHC neighbourhoods in BC") +
  theme_minimal() +
  theme(legend.position = "none", panel.grid.minor.x = element_blank())

```

To measure the impact of STRs on housing affordability in British Columbia, we developed a mixed-effect linear regression model predicting average rents as a function of the per-dwelling prevalence of dedicated short-term rentals. The model answers the question: "What is the predicted increase in average monthly rent in a community when the prevalence of dedicated STRs increases by one dedicated STR per 100 rental units?" We answer this question while controlling for year (since rents have been steadily increasing since 2016), community type, and the proportion of a neighbourhood's dwelling units which are renter-occupied. (Full methodological details are available in the appendix.)

Figure 10 shows the bivariate relationship between dedicated STRs per 100 rental units and average monthly rent for each community type. The patterns are noisy, but in each case the relationship is clear: neighbourhoods with a higher prevalence of commercial STRs have higher rents.

```{r fig_10, fig.width = 9}

find_outliers <- function(x) {
  q1 <- quantile(x, 0.25, na.rm = TRUE)
  q3 <- quantile(x, 0.75, na.rm = TRUE)
  iqr <- (q3 - q1) * 1.5
  which(x < q1 - iqr | x > q3 + iqr)
}

cmhc_str[-find_outliers(cmhc_str$iv), ] |> 
  mutate(tier = "All") |> 
  bind_rows(cmhc_str[-find_outliers(cmhc_str$iv), ]) |> 
  mutate(year = year + 2016) |> 
  filter(!is.na(tier)) |> 
  ggplot(aes(iv, total_rent)) +
  geom_point(aes(color = year), size = 0.5) +
  geom_smooth(color = "black", se = FALSE, method = "lm") +
  scale_y_continuous(name = "Average monthly rent", labels = scales::dollar) +
  scale_x_continuous(name = "Dedicated STRs per 100 rental units") +
  scale_color_viridis_b(option = "B") +
  ggtitle("Bivariate relationship between dedicated STRs and average monthly rent") +
  facet_wrap(~tier) +
  theme_minimal()

```

We formalize this finding in Table 1, which presents the mixed-effect regression model predicting monthly rents using the prevalence of dedicated STRs. The model accounts for nearly all of the observed variation in average monthly rents (r2 = 0.966), and confirms a strong independent effect of dedicated STRs on rents. Controlling for the other factors present in the model, an increase of one dedicated STR per 100 rental units predicts that average rents in the neighbourhood will be `r model_iv_coef_dollar` higher. (The model also finds that each year since 2016 is associated with a `r model_year_coef_dollar` increase in average rent, and that each additional percentage point of a neighbourhood's housing units which is composed of rental units is associated with a `r model_renter_coef_dollar` increase in average rent. Finally, the model offers different starting rents for each community type, assuming no commercial STRs, no renters, and the year 2016.)

```{r table_1, results = "asis"}

model_pretty <- model

names(model_pretty$coefficients) <- 
  c('Dedicated STRs / 100 rental units', 'Renter (%)', 'Year (0 = 2016)',
    'Tier - census agglomeration', 'Tier - central city', 
    'Tier - census metropolitan area', 'Tier - Resort/non-urban')

stargazer::stargazer(model_pretty, type = "latex", single.row = TRUE, 
                     dep.var.labels = "Average monthly rent")

```

### The burden of STRs on BC renter households

From 2016-2021, BC tenants in large and mid-sized cities paid `r rent_2016_2021_dollar` in rent. During the same time period, we estimate that `r rent_str_pct_2016_2021` (`r overpaid_2016_2021`) of this amount was due to the presence of commercial STRs. Put differently, if there had been no commercial STRs, BC renters would have saved `r overpaid_2016_2021` in rent. In 2019, when the STR market was at its pre-pandemic peak, we estimate dedicated STRs were responsible for `r rent_str_pct_2019` of the total rents paid by tenants. This is a province-wide average; in many communities, the amount of rent increases attributable to STRs was substantially higher.

Another way of framing these results is to look at the share of the *change* in rent which is attributable to the *change* in dedicated STR prevalence. The conclusion is sobering. During the 2017-2019 pre-pandemic period, the proliferation of dedicated STRs across BC accounts for fully `r str_incr_2017_2019` of the increases in rents which BC households were subjected to. The average BC neighbourhood in a medium or large city saw a `r rent_month_2017_2019` increase in monthly rent each of these years, of which `r str_incr_month_2017_2019` can be attributed to the growth of dedicated STRs. (As a plausibility check on these results, they are quite close to Barron et al. (2020)'s independent finding that approximately 20% of the rent increases in the 100 largest US cities from 2012-2016 were attributable to the growth of Airbnb.)

In 2020, when the number of commercial STRs plummeted because of the pandemic, the opposite effect occurred: the return of these units to the long-term market exerted downward pressure on rising rents. Our estimate is that 2020 rents increased `r str_incr_2020` less than they otherwise would have, because of the decrease in commercial STRs that year. The average BC neighbourhood in a medium or large city saw a `r rent_month_2020` increase in monthly rent, which was `r str_incr_month_2020` less than expected because of the decline in dedicated STRs.

Figure 11 and Table 2 summarize the estimated STR rent burden for each community type in BC. They make clear that an enormous proportion of the pre-pandemic increase in rent in smaller communities can be ascribed to the explosive growth of dedicated STRs in these areas. (Although the results in the "resort towns and non-urban" category should be taken as very broad estimates, since CMHC rental market data only covers a small portion of these communities province-wide, and therefore the results we have are not certain to generalize to the remaining such communities for which data is not available.) And likewise, an enormous cause of the much lower increase in average rents experienced by large central cities in 2020 can be ascribed to the collapse of those cities' STR markets during the pandemic, and the resulting return of many commercial STRs to the long-term rental market.

```{r fig_11}

rent_change_table_year |> 
  select(year, tier, med_rent, raw_rent) |> 
  pivot_longer(med_rent:raw_rent) |> 
  arrange(year, tier, desc(name)) |> 
  ggplot() +
  geom_col(aes(year, value, group = name, fill = name),
           position = position_dodge(width = -0.5)) +
  geom_segment(data = rent_change_table_year,
                mapping = aes(x = year, xend = year, y = raw_rent, 
                              yend = med_rent),
               size = 1.2,
               arrow = arrow(length = unit(0.3, "cm")),
               position = position_nudge(x = -0.125)) +
  scale_y_continuous(name = NULL, labels = scales::dollar) +
  scale_x_continuous((name = NULL)) +
  scale_fill_brewer(name = NULL, palette = "Accent", labels = c(
    "Actual median rent change",
    "Estimated median rent change with no change in dedicated STRs")) +
  facet_wrap(~tier) +
  theme_minimal() +
  theme(legend.position = "bottom")

```

```{r table_2}

rent_change_table |> 
  select(-starts_with("mean")) |> 
  pivot_wider(names_from = year, values_from = med_rent:str_incr) |> 
  relocate(ends_with("2019"), .after = tier) |> 
  mutate(across(c(med_rent_2017_2019, med_str_2017_2019, med_rent_2020,
                  med_str_2020), scales::dollar, 1),
         across(contains("incr"), scales::percent, 0.1)) |> 
  mutate(med_str_2017_2019 = paste0(med_str_2017_2019, " (", med_incr_2017_2019, ")"),
         med_str_2020 = paste0(med_str_2020, " (", med_incr_2020, ")")) |> 
  select(-med_incr_2017_2019, -med_incr_2020) |> 
set_names(c("Community type", 
            "Median YOY monthly rent chg. (2017-19)",
            "Median YOY impact of STR chg. on monthly rent chg. (2017-19)",
            "Total YOY impact of STR chg. on monthly rent chg. (2017-19)",
            "Median YOY monthly rent chg. (2020)",
            "Median YOY impact of STR chg. on monthly rent chg. (2020)",
            "Total YOY impact of STR chg. on monthly rent chg. (2020)")) |> 
  gt::gt()

```

### Trend analysis: STR rent burden

The sharp change in the impact of commercial STRs on rents in BC during the pandemic raises the question of what would have happened if commercial STRs had continued to grow on their pre-pandemic trajectory---and, more importantly, what will happen if they return to that trajectory after travel returns to normal. In the absence of significant new regulatory action, this is the most likely future scenario. Accordingly, Figure 12 extends the STR housing loss trend analysis from Figure 8, above, through the end of 2023. To be clear, it is impossible to make any concrete STR forecasts that far into the future, given the continuing uncertainty associated with the Covid pandemic, and the possibility of provincial or municipal regulatory action, but this analysis offers a plausible estimate of what a return to the pre-pandemic STR status quo would like in BC, and what the implications for renters would be.

```{r fig_12, fig.width = 9}

housing_loss_daily_model_2023 |> 
  group_by(tier) |>
  mutate(units_trend = slide_dbl(units_trend, mean, .before = 6, .complete = TRUE)) |>
  ungroup() |>
  select(tier, date, units, units_trend) |> 
  pivot_longer(-c(tier, date)) |> 
  filter(!is.na(value)) |> 
  mutate(label = case_when(
    tier == "All" & date == "2019-05-05" & name == "units" ~ "Actual housing loss", 
    tier == "All" & date == "2020-03-07" & name == "units_trend" ~ "Expected housing loss",
    TRUE ~ NA_character_)) |> 
  ggplot() +
  geom_ribbon(aes(x = date, ymin = units, ymax = units_trend, group = 1),
              data = {
                housing_loss_daily_model_2023 |> 
                  group_by(tier) |>
                  mutate(units_trend = slider::slide_dbl(
                    units_trend, mean, .before = 6, .complete = TRUE)) |>
                  ungroup()}, fill = col_palette[3], alpha = 0.3) +
  geom_line(aes(date, value, color = name), lwd = 1) +
  geom_label(aes(date, value, label = label, color = name), size = 3) +
  scale_x_date(name = NULL, limits = as.Date(c("2018-01-01", NA))) +
  scale_y_continuous(name = NULL, limits = c(0, NA), 
                     label = scales::comma) +
  scale_color_manual(name = NULL, 
                     labels = c("Actual STR housing loss", 
                                "Expected STR housing loss"), 
                     values = col_palette[c(5, 1)]) +
  facet_wrap(~tier, scales = "free_y") +
  theme_minimal() +
  theme(legend.position = "none",
        panel.grid.minor.x = element_blank(),
        plot.background = element_rect(fill = "white", colour = "transparent"))

```

If the province's STR market returns to its pre-pandemic trajectory by the end of 2023, this would imply that there will be `r housing_loss_2023` housing units operating as dedicated STRs---a `r housing_loss_change_2021_2023` increase over the actual number as of the end of 2021. This implies that STR activity will drive up average monthly rents in medium and large cities in the province by `r rent_inc_monthly_2021_2023` above their levels at the end of 2021. This is an extra `r rent_inc_annual_2021_2023` the average renter household will be paying per year because of the growth of commercial STRs---`r rent_total_2021_2023` in total each year.
