---
title: "Chapter 4. STR regulatory pathways"
date: "31/05/2022"
output: pdf_document
---

```{r setup, include = FALSE}

knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)

library(tidyverse)
library(slider)

qs::qload("output/data/ch_4.qsm")

col_palette <-
  c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

```

A growing number of municipalities and provinces both in Canada and globally have started enacting regulations around the operation of short-term rentals.

TKTK The governments can use multiple methods for going about regulating STRs. One of the most crucial elements for a successful regulation of STRs is for governments to make a distinction between home sharing and commercially-operated STRs. Home sharing should be encouraged while commercial STRs should be made illegal. The evidence suggests that the most effective way to regulate STRs is through principal-residence requirements with mandatory registration. All households wishful to operate a short-term rental could do so in their residence with a valid and unique registration number. We present two stories of municipalities having regulated STRs -- Vancouver and Toronto -- and also present the Province of Quebec's law project that aims to regulate STRs at the provincial level.

In this final chapter we examine several recent Canadian attempts to regulate STRs, and conclude by offering a set of principles to guide the construction of a provincial STR regulation framework in British Columbia---a task that is now overdue.

# Vancouver: The early success story

In April 2018, the City of Vancouver enacted regulations on the operations of short-term rentals in the City, defined as rentals offered for thirty or fewer consecutive days (City of Vancouver, 2020a). Under these regulations, each STR operator is required to obtain a license for their rental unit, the license being valid for one year. The listing can only be operated out of the host's principal residence, either for the entire dwelling or for individual rooms. The registration is mandatory on any rental platform, although Airbnb is the only platform that agreed to require hosts in Vancouver to fill out a license field in their online listing, to engage in data sharing, and to undertake operator education (City of Vancouver, 2020b). In August 2018, shortly before the City's announced start date for enforcement of the registration system, Airbnb removed approximately 2,400 listings which had not received licenses, which amounted to 30.8% of all displayed listings. Using trend analysis, we estimated that, by mid-2019, the Vancouver STR regulation reduced active STR listings by more than one third, compared to the counter-factual scenario where the regulations had not been introduced. We also estimated that the City's regulations reduced commercial STRs by a similar proportion, thus returning more than 800 housing units to the long-term market. The analysis in chapter 3 of this report demonstrates that these regulatory actions have exerted significant downward pressure on the growth of residential rents in Vancouver.

Further evidence supporting the notion that Vancouver's STR regulations have been effective at reducing the dedicated STRs which drive up housing costs comes from comparing Vancouver with peer cities nationally and with the rest of the Metro Vancouver region. As Figure X demonstrates, Vancouver's STR growth trajectory lagged significantly behind Montreal and Toronto (which did not have comparable STR restrictions at the time, and thus serve as useful contrasting cases) from 2018 through the end of 2019, and the difference was particularly large with respect to frequently-rented entire home (FREH) listings. Through 2017, FREH listings were growing at nearly identical rates in Toronto and Vancouver, while by 2019 Toronto had nearly twice the total FREH growth which Vancouver had. Figure X demonstrates that the same pattern holds true comparing Vancouver to the rest of the CMA---the growth rate of FREH listings in the rest of the CMA has vastly exceeded Vancouver's since 2018.

```{r fig_4_1}

canada_comparison |> 
    filter(date >= "2016-07-01", date <= "2019-12-31") %>% 
  group_by(city, type) %>% 
  mutate(index = 100 * n / n[date == "2017-01-01"]) %>% 
  ungroup() %>% 
  mutate(type = case_when(
    type == "active" ~ "Active listings", 
    type == "FREH" ~ "FREH listings",
    type == "multi" ~ "Multilistings")) %>% 
  ggplot(aes(date, index, colour = city, size = city)) +
  geom_line() +
  scale_y_continuous(name = NULL, label = scales::comma) +
  scale_x_date(name = NULL, limits = c(as.Date("2016-07-01"), NA)) +
  scale_colour_manual(name = NULL, 
                      values = col_palette[c(3, 6, 5)],
                      guide = guide_legend(override.aes = list(
                        size = c(0.5, 0.5, 1)))) +
  scale_size_manual(values = c("Vancouver" = 1, "Montreal" = 0.5,
                               "Toronto" = 0.5), guide = "none") + 
  facet_wrap(vars(type), nrow = 1) +
  theme_minimal() +
  theme(legend.position = "bottom", 
        panel.grid.minor.x = element_blank())

```

```{r fig_4_2}

cma_comparison |> 
    filter(date >= "2016-07-01", date <= "2019-12-31") |> 
  group_by(city, type) |> 
  mutate(index = 100 * n / n[date == "2017-01-01"]) |> 
  ungroup() |> 
  mutate(type = case_when(
    type == "active" ~ "Active listings", 
    type == "FREH" ~ "FREH listings",
    type == "multi" ~ "Multilistings")) |> 
  ggplot(aes(date, index, colour = city, size = city)) +
  geom_line() +
  scale_y_continuous(name = NULL, label = scales::comma) +
  scale_x_date(name = NULL, limits = c(as.Date("2016-07-01"), NA)) +
  scale_colour_manual(name = NULL, 
                      values = col_palette[c(3, 5)],
                      guide = guide_legend(override.aes = list(
                        size = c(0.5, 1)))) +
  scale_size_manual(values = c("Vancouver" = 1, "Rest of CMA" = 0.5), 
                    guide = "none") + 
  facet_wrap(vars(type), nrow = 1) +
  theme_minimal() +
  theme(legend.position = "bottom", 
        panel.grid.minor.x = element_blank())

```

# Toronto: The importance of closing loopholes

In early 2021, the City of Toronto's STR regulations came into force, requiring all STR listings to be registered, and limiting STRs to a host's principal residence. The vast majority of hosts failed to register their listings, however. In theory, this means that they should have been removed from online STR platforms such as Airbnb. And indeed, just prior to the implementation date in early January, Airbnb removed 2,600 Toronto listings from its site. However, this was only a small fraction of the listings which were not registered, and Airbnb forcibly shifted the overwhelming majority of remaining listings to a minimum rental length of 28 days, which exempts them from the registration requirement.

According to an analysis we conducted for the City of Toronto (Wachsmuth et al. 2021b), in November of 2020, only 6.3% of Toronto listings had a minimum stay of 28 days or more, while at the beginning of February 2021 it was 72.2%. While it was too early to tell at the time of our analysis what the long-term impacts of Toronto's STR regulations would be, it is clear that the fact that STR with a minimum 28-day reservation period are exempted from the need to register has created a serious loophole in Toronto's registration scheme. we can see that offering an option for hosts not to register acts as a serious barrier to effective regulation. Toronto's registration system was successful at shifting market behavior, but by diverting properties out of the system, instead of capturing them within the system. Effective STR rules, therefore, should recognize the 28-day-minimum loophole and work to close it.

# Quebec: A provincial registration system

TKTK To promote efficiency and prevent municipalities shifting the burden, the province should weigh in and establish a single registration system. This form of regulation is being considered by the Province of Quebec. In August 2021, a law project, Le projet de loi 100, was submitted to Quebec's National Assembly. These proposed regulations would make mandatory the registration of all short-term accommodations in the province, including but not limited to short-term rentals, hotels, camping sites, and B&Bs. While the proposed law is a good model, it is clear from the enforcement troubles experienced by municipalities attempting to regulate STRs that some additional elements should be considered.

TKTK Indeed, it is essential that a registration system should make the platforms accountable, by allowing only registered listings to be displayed on their site. The experience of other jurisdictions shows that any conditions that allow short-term rental hosts to avoid registration will be subject to enormous pressure to provide a loophole for bad actors. The province is in a strong position to launch and enact the registration system.

TKTK Another crucial element of every STR regulation is the regulation's capacity to differentiate between home sharing (i.e. principal residence) and commercial operations. The negative impacts of STRs, namely housing unaffordability, nuisance, and crime tend to be disproportionately distributed among commercially-operated STRs. The regulation of short-term rentals should not systematically encourage or discourage short-term rentals in relation to hotels, since the overall benefits are relatively similar. But there is a strong public policy case, particularly in cities, for regulations to systematically discourage commercial short-term rentals relative to casual short-term rentals, since the former are responsible for most of the negative impacts. Given the negative impact of STRs across British Columbia, the default approach of all municipalities should be to consider principal residence STRs only.

# Regulatory principles for BC

As British Columbia emerges from the Covid pandemic, it is likely that the STR market will quickly return to its pre-pandemic status quo. This means that the temporary respite from STR-induced housing loss and rising rents will likely reverse itself. Meanwhile, regulation of STRs is piecemeal at best, and many communities likely to bear the brunt of the housing cost impacts do not have adequate frameworks in place to mitigate these impacts. Now is thus an excellent time for BC's STR regulations to be updated, and the simplest, most effective approach is for the Province should take the lead by establishing a single mandatory STR registration system for all of BC, while municipalities strongly consider establishing principal-residence requirements to redirect their STR markets towards home sharing and away from commercial operations.

It is helpful to distinguish between three tasks that regulators need to accomplish: 1. They must decide on the appropriate regulations for the jurisdiction. 2. they must gather the information necessary to enforce the rules 3. they must enforce the rules. Municipalities are the best positioned to tackle the first task: deciding on the most appropriate rules. Different communities can and should have different priorities for how they choose to balance STRs and housing issues. Larger cities generally have better substitutes for STRs (hotels) and have more severe housing affordability issues. Rural communities may not have good substitutes for STRs, may be more economically reliant on seasonal tourism, and may have somewhat less severe housing issues. So there will not be a viable "one size fits all" model for the province. However, the severity and ubiqiutiy of housing affordability issues in British Columbia suggests that many if not all municipalities should strongly consider introducing a principal-residence requirement into their STR rules. Empirical research, in addition to the pioneering experience of the City of Vancouver, strongly suggests that a principal residence requirement is the best means of redirecting STR activity away from commercial operations and toward home sharing, while simultaneously reducing the housing affordability burden of STRs on local residents.

By contrast, with respect to the second task, information gathering, there very much is a viable "one size fits all" model, and it is one in which the Province takes the lead. The provincial government should follow the lead of Quebec and establish a single mandatory registration system for all STRs operating in British Columbia. The information requirements for a registration system do not vary much between localities, so there are massive economies of scale in having a single system. Large municipalities are probably in a position to follow Vancouver's lead and implement registration systems on their own, but for the vast majority of BC municipalities, this isn't a viable possibility. The experience of Airbnb shifting listings to a 28-day minimum in Toronto to avoid the need to comply with regulations further implies that the Province should not make a distinction between "short-term" and "long-term" rentals for the purposes of registration. Instead, the rules could identify a set of online STR platforms, and require the registration of any listing which is advertised on these platforms, regardless of the length of reservation.

Finally, regulatory enforcement should be a responsibility shared between both the Province and the municipalities. The Province can enforce the use and validity of the registration system, while municipalities can use the information in the registration system to enforce local bylaws. They can use their local knowledge, gained from inspections and complaints, to report registration problems to the Province, with the result that the overall STR regulatory system could be self-strengthening.

Establishing such a system will by no means solve all of BC's housing affordability problems. But, compared with the longer-term solutions which will be needed to ensure that the province's housing supply is adequate to housing needs, better regulation of short-term rentals is arguably the lowest hanging fruit capable of meaningfully addressing rapidly escalating housing costs.
